---
title: "Data Cleaning Script"
author: "Rebecca Folmer Schade & Sophia Kleist Karlson"
date: "2021-05-11 updated 2021-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
```

## Loading packages 
```{r}
# Quick install/load of all used packages
# install.packages("pacman")
# pacman::p_load(pacman, osmdata, sf, tidyverse, raster)
library(osmdata)
library(sf)
library(tidyverse)
library(raster)
```
# Data gathering and cleaning 
In this script, data on LGBT+ friendly amenities in the European Economic Area (EEA) and Switzerland is gathered from Open Street Maps and cleaned for use in the Leaflet visualization.

## Querying data from Open street maps
### Creating bounding boxes
Because of the scattering of European territories, the data will be queried within three bounding boxes - one containing most of Europe, one containing the Danish territory Greenland, and one containing the French territory French Guiana.

```{r}
# Bounding box for most of Europe
box = c(xmin = -28.3, ymin = 31.1, xmax = 45.7, ymax = 71.7)
# Bounding box for Greenland 
greenland_box <- c(xmin = -76.6, ymin = 57.3, xmax = -9.0, ymax = 84.0)
# Bounging box for French Guiana
guiana_box <- c(xmin = -63.19, ymin = -0.33, xmax = -50.13, ymax = 9.59)
```

### Querying the data
Data is queried withing the three bounding boxes that contain the keys gay= and lgbtq= with all associated values indicating that the location is lgbt+ friendly (gay=yes/welcome/only and lgbtq=primary/only/yes). Though the wiki for the lgbtq key discourages the use of lgbtq=yes, it still queried as some data was coded as such [(link here, last read 2021-05-25)](https://wiki.openstreetmap.org/wiki/Key:lgbtq).  

*Europe (with the exception of Greenland and French Guiana)*  

```{r}
# Turning 'box' into a bbox object
EEA_bb  <- st_bbox(box)
class(EEA_bb)
# Defining the bbox for the OPenStreetMap request
EEA_q <- opq(bbox = EEA_bb, timeout = 180)
# Defining requests 
EEA_qa <- add_osm_feature(EEA_q, key = 'lgbtq',value = 'primary')
EEA_qb <- add_osm_feature(EEA_q, key = 'lgbtq',value = 'only')
EEA_qc <- add_osm_feature(EEA_q, key = 'lgbtq',value = 'yes')
EEA_qd <- add_osm_feature(EEA_q, key = 'gay',value = 'yes')
EEA_qe <- add_osm_feature(EEA_q, key = 'gay',value = 'welcome')
EEA_qf <- add_osm_feature(EEA_q, key = 'gay',value = 'only')
# Requesting data 
# (as querying the data all at once seemed to cause trouble with the limit of requests for one's IP adress, data associated with different keys and values is queried separately)
# (This is time consuming, but at least it works if you wait a little inbetween requests)
lgbtq_primary <- osmdata_sf(EEA_qa)
lgbtq_only <- osmdata_sf(EEA_qb)
lgbtq_yes <- osmdata_sf(EEA_qc)
gay_yes <- osmdata_sf(EEA_qd)
gay_welcome <- osmdata_sf(EEA_qe)
gay_only <- osmdata_sf(EEA_qf)
# Checking how many datapoints were found
unique_osmdata(lgbtq_primary)
unique_osmdata(lgbtq_only)
unique_osmdata(lgbtq_yes)
unique_osmdata(gay_yes)
unique_osmdata(gay_welcome)
unique_osmdata(gay_only) 

# Gathering the data 
EEA_safe_spaces <- c(lgbtq_primary,
                     lgbtq_only, 
                     lgbtq_yes,
                     gay_yes, 
                     gay_welcome)
# Having a closer look at the EEA_safe_spaces object
class(EEA_safe_spaces)
unique_osmdata(EEA_safe_spaces)
# > The object contains not only points and polygons, but also lines and multipolygons, which is unexpected. 
```

*Greenland*  

```{r}
# Turning 'greenland_box' into a bbox object
greenland_bb  <- st_bbox(greenland_box)
class(greenland_bb)
# Defining the bbox for the OPenStreetMap request
greenland_q <- opq(bbox = greenland_bb, timeout = 180)
# Defining requests 
greenland_qa <- add_osm_feature(greenland_q, key = 'lgbtq',value = 'primary')
greenland_qb <- add_osm_feature(greenland_q, key = 'lgbtq',value = 'only')
greenland_qc <- add_osm_feature(greenland_q, key = 'lgbtq',value = 'yes')
greenland_qd <- add_osm_feature(greenland_q, key = 'gay',value = 'yes')
greenland_qe <- add_osm_feature(greenland_q, key = 'gay',value = 'welcome')
greenland_qf <- add_osm_feature(greenland_q, key = 'gay',value = 'only')
# Requesting data 
greenland_lgbtq_primary <- osmdata_sf(greenland_qa)
greenland_lgbtq_only <- osmdata_sf(greenland_qb)
greenland_lgbtq_yes <- osmdata_sf(greenland_qc)
greenland_gay_yes <- osmdata_sf(greenland_qd)
greenland_gay_welcome <- osmdata_sf(greenland_qe)
greenland_gay_only <- osmdata_sf(greenland_qf)
# Checking how many datapoints were found
unique_osmdata(greenland_lgbtq_primary)
unique_osmdata(greenland_lgbtq_only)  
unique_osmdata(greenland_lgbtq_yes) 
unique_osmdata(greenland_gay_yes)  
unique_osmdata(greenland_gay_welcome)
unique_osmdata(greenland_gay_only) 
# > No datapoints were found
```

*French Guiana*

```{r}
# Turning 'guiana_box' into a bbox object
guiana_bb  <- st_bbox(guiana_box)
class(guiana_bb)
# Defining the bbox for the OPenStreetMap request
guiana_q <- opq(bbox = guiana_bb, timeout = 180)
# Defining requests 
guiana_qa <- add_osm_feature(guiana_q, key = 'lgbtq',value = 'primary')
guiana_qb <- add_osm_feature(guiana_q, key = 'lgbtq',value = 'only')
guiana_qc <- add_osm_feature(guiana_q, key = 'lgbtq',value = 'yes')
guiana_qd <- add_osm_feature(guiana_q, key = 'gay',value = 'yes')
guiana_qe <- add_osm_feature(guiana_q, key = 'gay',value = 'welcome')
guiana_qf <- add_osm_feature(guiana_q, key = 'gay',value = 'only')
# Requesting data 
guiana_lgbtq_primary <- osmdata_sf(guiana_qa)
guiana_lgbtq_only <- osmdata_sf(guiana_qb)
guiana_lgbtq_yes <- osmdata_sf(guiana_qc)
guiana_gay_yes <- osmdata_sf(guiana_qd)
guiana_gay_welcome <- osmdata_sf(guiana_qe)
guiana_gay_only <- osmdata_sf(guiana_qf)
# Checking how many datapoints were found
unique_osmdata(guiana_lgbtq_primary)
unique_osmdata(guiana_lgbtq_only) 
unique_osmdata(guiana_lgbtq_yes) 
unique_osmdata(guiana_gay_yes)
unique_osmdata(guiana_gay_welcome) 
unique_osmdata(guiana_gay_only)
# > No datapoints were found
```

## Cleaning the data
### Creating a map of countries in the EEA (+ Switzerland)
While the main visualizations are made in leaflet, this is useful for quick illustrations and subsetting data. 
The map is made with data from © EuroGeographics for the administrative boundaries. 

```{r}
# loading data from eurostat 
world_geo <- st_read("data/CNTR_RG_01M_2020_3035.shp/CNTR_RG_01M_2020_3035.shp")
head(world_geo)

# Let's try plotting the data 
plot(st_geometry(euro_geo))

# Re-projecting map to EPSG:4668 - a 2D map is easier for visualization purposes
world_transform <- st_transform(world_geo, crs = 4668)

# Let's try plotting the data - again
plot(st_geometry(world_transform))

# Creating a list of countries that are members of the EEA 
EEA_members <- c("Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czechia", "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary", "Iceland", "Ireland", "Italy", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Malta", "Netherlands", "Norway", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Sweden", "Switzerland")
# > With the excption of Switzerland, the list was taken from https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Glossary:European_Economic_Area_(EEA) (2021-05-25)

# Checking to see if Anything was overlooked in this list 
sort(unique(world_transform$NAME_ENGL))
# > Though a part of the Danish Commonwealth, Greenland and the Faroes have their own entries in the dataset 

# Adding the Faroes to the list 
EEA_members <- c(EEA_members, "Faroes", "Greenland")
EEA_members

# Checking that all countries in the EEA_members list are present in the dataset 
EEA_members %in% world_transform$NAME_ENGL

# Creating a map only with the countries in the EEA for subsetting data later on
EEA_geo <- world_transform %>% dplyr::filter(NAME_ENGL %in% EEA_members)

# Showing EEA territories in a plot 
plot(st_geometry(world_transform));plot(st_geometry(EEA_geo), col = "red", fil = "red", add = TRUE)

# Setting the bounding box to one more focused on Europe for quick visualizations
attr(st_geometry(world_transform), "bbox") <-  st_bbox(box)
plot(st_geometry(world_transform))

# Create a bounding box polygon to illustrate the bounding box
graph_box <- st_make_grid(box, n = 1)
plot(st_geometry(world_transform));plot(graph_box, add = TRUE)
```


### Getting rid of duplicated datapoints and those outside the area of interest 

```{r}
# Removing duplicated geometries
EEA_safe_unique <- unique_osmdata(EEA_safe_spaces)

# Isolate the points and change projection
EEA_rpoint <- EEA_safe_unique$osm_points %>%
  st_transform(crs = 4668) 

# Isolate the polygons Compute centroids of the polygons and store data as points 
EEA_rpoly  <- EEA_safe_unique$osm_polygons %>%
  st_transform(crs = 4668) %>% 
  st_centroid()

# > I am saving the lines and multipolygons in separate objects because I don't know what to do with them, but I don't want to have to query them again in case I could use them. 
EEA_rline <- EEA_safe_unique$osm_lines %>% 
  st_transform(crs = 4668)
EEA_rmultipoly <- EEA_safe_unique$osm_multipolygons %>% 
  st_transform(crs = 4668)

# Merging
# Keeping the columns that are present in all dataframes
EEA_keep <- which(colnames(EEA_rpoint) %in% colnames(EEA_rpoly)[colnames(EEA_rpoly) %in% colnames(EEA_rpoint)])
EEA_keep
EEA_keep_names <- colnames(EEA_rpoint)[EEA_keep]
EEA_keep_names
EEA_rpoint_selected <- EEA_rpoint %>% dplyr::select(all_of(EEA_keep_names))
EEA_rpoly_selected <- EEA_rpoly %>% dplyr::select(all_of(EEA_keep_names))
EEA_safe_osm <- rbind(EEA_rpoly_selected,EEA_rpoint_selected)

# Remove points outside the area of interest
EEA_safe_osm_filtered <- st_intersection(EEA_safe_osm, (EEA_geo %>% st_geometry() %>% st_union()))

# Plotting data 
plot(st_geometry(world_transform));plot(st_geometry(EEA_safe_osm_filtered),col = "red", add = TRUE)
```

### Having a look at the lines and multipolygons 
Neither the lines nor the multipolygons show up on a plot, so for not I have elected to exclude them.

```{r}
plot(st_geometry(world_transform));plot(st_geometry(EEA_multipoly_select),col = "red", add = TRUE)
plot(st_geometry(EEA_line_select))
EEA_line_select
world_transform
```

### Selecting important variables
Not all of the variables in the dataset from OSM are useful for our visualization - here we remove the irrelevant ones. 

```{r}
# Removing unneeded variables
names(EEA_safe_osm_filtered)
EEA_point_select <- EEA_safe_osm_filtered %>% dplyr::select(osm_id, name, amenity, opening_hours, website)
names(EEA_rline)
EEA_line_select <- EEA_rline %>% dplyr::select(osm_id, name, amenity, opening_hours, website)
names(EEA_rmultipoly)
EEA_multipoly_select <- EEA_rmultipoly %>% dplyr::select(osm_id, name, amenity)
```

### Cleaning variables 
As these variables will be used in the visualization, they should be made a little prettier to look at. 

```{r}
#=====> Cleaning EEA_point_select <=====#
#===> Cleaning $name
unique(EEA_point_select$name)
# > I identified three issues: The translation of special characters, some entries with small first letters, and the use of quetation marks. 
# Fixing the encoding of local characters to UTF-8
EEA_point_select$name <- iconv(EEA_point_select$name, from = "UTF-8", to = "UTF-8")
# Fixing the small first letters 
EEA_point_select$name <- str_to_title(EEA_point_select$name)
# Fixing the unneded quotation marks
EEA_point_select$name <- str_replace_all(EEA_point_select$name, pattern = "\"", replacement = "")

#===> Cleaning $amenity
unique(EEA_point_select$amenity)
# > two issues: the use of underscores instead of space and lack of capitalization
# Fixing the use of underscores 
EEA_point_select$amenity <- str_replace_all(EEA_point_select$amenity, pattern = "_", replacement = " ")
# Fixing the capitalization 
EEA_point_select$amenity <- str_to_sentence(EEA_point_select$amenity)

#===> Cleaning $opening_hours
unique(EEA_point_select$opening_hours)
# > One immidiate issue identified: Encoding of characters
# Fixing encoding 
EEA_point_select$opening_hours <- iconv(EEA_point_select$opening_hours, from = "UTF-8", to = "UTF-8")

#===> Cleaning $website
unique(EEA_point_select$website)
# > Probably best not to mess with this - there does not seem to be any obvious issues
```
## Saving data 
As the lines and multypoligons were unexpected, they are saved separately for now.

```{r}
# Saving data in the data folder 
st_write(EEA_point_select, "data/EEA_points.shp")
st_write(EEA_line_select, "data/EEA_lines.shp")
st_write(EEA_multipoly_select, "data/EEA_multypolygons.shp")
# Checking if it worked 
test_df <- read_sf("data/EEA_points.shp")
# >  It did :-) 
```
